---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
The data used can be downloaded [here](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)  
This is a .zip file containing the .csv file used in this study.  

## Loading and preprocessing the data
```{r echo = TRUE}
options(scipen = 2, digits = 2)
```

```{r cache=TRUE,echo = TRUE}
options(scipen = 2, digits = 2)
rawdata<-read.csv("activity.csv")
data<-rawdata
data$date<-as.Date(data$date)

```

## What is mean total number of steps taken per day?
```{r echo = TRUE}
data<-data[!is.na(data$steps),]
dates<-unique(data$date)
daysteps<-numeric()
for(i in 1:length(dates)){
  daysteps[i]<-sum(data$steps[data$date==dates[i]])
}
hist(daysteps,main="Histogram - Steps per day",xlab="Steps per Day",col="steelblue")
meansteps<-mean(daysteps)
mediansteps<-median(daysteps)
```
I am ignoring the dates that have no data and removing those dates from the dataset  
The mean (average number of steps per day) is **`r meansteps`** and the median is **`r mediansteps`**


## What is the average daily activity pattern?
```{r echo = TRUE}
intervals<-unique(data$interval)
dailyPattern<-tapply(data$steps,data$interval,mean)
plot(intervals,dailyPattern,type="l",ylab="Average number of Steps",xlab="Interval",main="Daily Activity Pattern")
max_index<-which.max(dailyPattern)
max_interval<-intervals[max_index]
max_steps<-dailyPattern[max_index]
```
The 5 minutes interval **`r max_interval`** contains the maximum number of steps in average with a value of `r max_steps` steps.


## Imputing missing values
```{r echo = TRUE}
data<-rawdata
data$date<-as.Date(data$date)
missingValues<-sum(is.na(data$steps))
missingValueIndex<-which(is.na(data$steps))
for(i in 1:missingValues){
  data$steps[missingValueIndex[i]]<-dailyPattern[as.character(data$interval[missingValueIndex[i]])]
}
```
There are **`r missingValues`** missing values on this dataset  
Each missing value is replaced with the average number of steps made per day on that particular 5 minute interval accross all days.
```{r echo = TRUE}
daysteps<-numeric()
dates<-unique(data$date)
for(i in 1:length(dates)){
  daysteps[i]<-sum(data$steps[data$date==dates[i]])
}
hist(daysteps,main="Histogram - Steps per day",xlab="Steps per Day",col="steelblue")
meansteps<-mean(daysteps)
mediansteps<-median(daysteps)
```
Now the mean (average number of steps per day) is **`r meansteps`** and the median is **`r mediansteps`**  
The mean value doesn't change because I am using the mean value to fill the missing values. It can be expected that the median changes a little because the dataset has changed. In fact in this case the median takes the value of one of the added values.
Evidently inputing missing data into the dataset changes things, but since the criteria is not aggressive, the difference is very small

## Are there differences in activity patterns between weekdays and weekends?
```{r echo = TRUE}
library(ggplot2)
Sys.setlocale("LC_TIME", "English")
day<-numeric()
for(i in 1:length(data$date)){
  if(weekdays(data$date[i])=="Saturday" || weekdays(data$date[i])=="Sunday")
    day[i]<-1
  else
    day[i]<-0
}

day<-factor(day,labels=c("weekday","weekend"))
data[,4]<-day
names(data)[4]<-"day"
ggplot(data,aes(interval,steps))+facet_grid(.~day)+ stat_summary(fun.y="mean", geom="line")
```
